EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-06 03_edipack2.tex
GB pass on `sparse map` subsection (3.5.2 currently)
I don't like too much the formatting of the code snippets. But this is a wider issue that I will discuss with Adriano later.

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a general quantum impurity problem defined by the
following Hamiltonian:
$$
\hat{H} = \hat{H}_{imp} + \hat{H}_{bath} + \hat{H}_{hyb} + \hat{H}_{ph} + \hat{H}_{e-ph}
$$
which describes a  multi-orbital interacting quantum
impurity coupled to an electronic bath and to local, i.e. Holstein,
phonons. 
\begin{equation}\label{Himp}
  \begin{split}
    \hat{H}_{imp} & = \hat{H}^0_{imp} + \hat{H}^{int}_{imp}\\
    \hat{H}^0_{imp} & =
    \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}d_{\b\sigma'}\\
\end{split}
\end{equation}

@@

The bath part and its coupling to the impurity has the form: 
\begin{equation}\label{Hbath}
  \begin{split}
    \hat{H}_{bath} &=
    \sum_p\sum_{\a\b\sigma\sigma'}h^p_{\a\b\sigma\sigma'}a^{+}_{p\a\sigma}a_{p\b\sigma'}\\
    %
    \hat{H}_{hyb} &= \sum_p\sum_{\a\b\sigma\sigma'}V^p_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}a_{p\b\sigma'}+H.c. \\
\end{split}
\end{equation}
where $p=1,\dots,N_{bath}$ is an index running over a finite number of
bath elements, $a_{p\alpha\sigma}$ ($a^+_{p\alpha\sigma}$) are the destruction (creation) operators for
the bath electrons with index $p$, with orbital $\alpha$ and
spin $\sigma$.
The  properties of each bath level are described by the
matrices $h^p_{\a\b\sigma\sigma'}$. As such any bath element can be
composed of several electronic levels according the bath topology,
which will be discussed further in the following.
Each bath level couples to the impurity with an amplitude
$V^p_{\a\b\s\sigma'}$ which we allow to couple different orbital and
opposite spins.

Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, $b_q$
($b_q^+$) are the destruction (creation) operators for the phonon $q$
with frequency $\omega_{0q}$. The matrix  $g_{\a\b}$ expresses is the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$. 
In the following we consider a bath discretized into a  number
of bath degrees of freedom and a finite number of available phonons $N_{ph}$,
to cut-off the unbounded dimensions of the local phonons Hilbert space.

\paragraph{{\bf Remark}} As extensively discussed in
\onlinecite{Amaricci2022}, the inclusion of local phonons with
truncated dimensions ultimately amounts to a sequential application of
the procedures defined for the electronic part, i.e. one per phonon
mode. This includes a linear scaling of the dimensions with $N_{ph}$
and of course largely limit the available degrees of freedom to
describe electronic states. In the rest of the paper we focus
specifically on the electronic part of the quantum impurity problem,
referring to the phonons in specific sections when their presence
introduces non-trivial modifications.

More specifically, we consider a system composed of $N_{imp}=1$
impurities, i.e. a single impurity problem, $N_{bath}$ bath elements
and $N_{ph}$ phonons. The size of the system is determined by
the number of phonons (fixed) and that of {\it electronic} levels, i.e. levels with a local
electronic Hilbert space
$\HH_e=\{\ket{0},\ket{\up},\ket{\dw},\ket{\up\dw} \}$. 
Due to is internal structure, the single impurity contains
$N_i= N_\a$ electronic levels.
The total number of levels is then determined by
the electronic levels in the bath $N_b$. This is a function of the bath topology and $N_{bath}$,
i.e. the number of bath elements. In the simplest case each bath
element corresponds to an independent electronic levels coupled to the
impurity, thus $N_b\equiv N_{bath}$. 
We indicate with $N_s=N_i + N_b$ the total number of electronic levels.

The setup of the general quantum impurity problem is implemented in
different parts of the \NAME software. The dimensions of the system
are controlled by input variables {\tt Nspin}$=N_\sigma$,
{\tt Norb}$=N_\a$ and {\tt Nbath}$=N_{bath}$ in {\tt
  ED\_INPUT\_VARS}. These are used to
determine the variables {\tt Ns}$=N_s$ and $N_b$, defined in the global
memory pool {\tt ED\_VARS\_GLOBAL}, using the functions contained
in {\tt ED\_SETUP}. 
The user can define the local non-interacting Hamiltonian
$h^0_{\a\b\s\s'}$ using the function {\tt ed\_set\_hloc} in {\tt ED\_AUX\_FUNX}.
The matrix is then stored in the internal memory and shared throughout the
code.
On the other hand the setup of the bath matrices $h^p_{\a\b\s\s'}$
requires a more involved procedure which will be illustrated in
\secu{sSecBath}.

\subsection{The Fock basis states}\label{sSecBasis}
The Fock space of the quantum impurity problems is defined as
$\FF=\FF_e\otimes \FF_{ph}$, with $\FF_e=\bigoplus_{n=0}^{N_s}
S_-\HH_e^{\otimes n}$ the electronic Fock space,  $\FF_{ph}=\bigoplus_{n=0}^{N_q}S_+\HH_{ph}^{\otimes n}$ the
phonon Fock space,  $\HH_{ph}=\{\ket{0},\ket{1},\dots,\ket{N_{ph}}\}$ is the local phonon Hilbert space
and ($S_-$)  $S_+$ the (anti-)symmetrization operator.  
The total dimension of the Fock space is
$D=D_e\cdot D_{ph}=4^{N_s}\cdot (N_{ph}+1)$ making the exponential
growth with the number of electron levels transparent. 
The quantum states in the space $\FF$ are naturally represented in
terms of occupation number formalism of the second quantization,
i.e. the Fock basis.
For a system of $N_s$ electrons each Fock state
reads $\ket{p}\ket{\vec{n}}$ with
$$
\ket{\vec{n}}=\ket{\vec{n}_\up}{\vec{n}_\dw}=\ket{n_{1\up},\dots,n_{N_s\up},n_{1\dw},\dots,n_{N_s\dw}}
$$ 
where $p=1,\dots,N_{ph}$ is the number of local phonons while $n_{a\sigma}=0,1$ signals the absence or the
presence of an electron with spin $\sigma$ at the level $a$.
The electronic part of the Fock state $\ket{\vec{n}}$ is represented as a string of
zeros and ones of length $2N_s$. Thus, any such state can  be encoded
in a computer using a sequence of $2N_s$ bits or, analogously, as a
given integer $I=0,\dots 2^{2N_s}-1$ so that $\ket{\vec{n}}=\ket{I}$.  
Together with the basis states one defines  destruction and creation 
operators, respectively $c_{a\s}$ and $c^+_{a\s}$, which acts on the
Fock space as: 
$\ket{\vec{n}}$ as:  
\begin{align*}
  c_{a\sigma}\ket{\vec{n}} &=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};\qquad
    c^{+}_{a\sigma}\ket{\vec{n}} &=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
\end{align*}
where $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$ takes care
of the fermionic sign imposed by Pauli principle.

@@

\begin{table}%[ht]
  \label{TabSector}
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_s}{N_\up}\binom{N_s}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_s-S_z-2m}\binom{N_s}{N_s-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv[N_\up+N_\dw$ & $\binom{2N_s}{N}$ \\ 
 \hline
\end{tabularx}
\end{center}
\caption{Ciao}
\end{table}

The {\bf normal} case requires a brief remark. Because of the
independent conservation of $N_\up$ and $N_\dw$, the local Hilbert
space and the electronic Fock can be factorizes as, respectively,
$\HH=\HH_\up\otimes\HH_\dw$, $\FF_e = \FF_{e\up}\otimes \FF_{e\dw}$.  
Accordingly any Fock state is written as $\ket{\vec{n}_\up}\ket{\vec{n}_\dw}$, the symmetry sector can be written as  $\SS_{\vec{Q}} = \SS_{N_\up}\otimes
\SS_{N_\dw}$ and, correspondingly, the sector map splits in two
mutually exclusive ones $\MM = \MM_\up
\otimes \MM_\dw$.
Thus, each state $\ket{i}=\ket{i_\up}\ket{i_\dw}$ of the
sector is labelled by two integers $[i_\up,i_\dw]$, 
$i_\sigma=1,\dots,D_{\SS_\sigma}$ such that $i=i_\up + i_\dw
D_{\SS_\dw}$. The map $\MM$ connects any such state to a Fock state
$\ket{I}=\ket{I_\up}\ket{I_\dw}$ labelled by two integers
$[I_\up,I_\dw]$ as $I=I_\up +   I_\dw 2^{N_s}$. For a thorough
discussion about Fock basis organization in this case see
\onlinecite{Amaricci2022}.

@@

where {\tt NORB} is the numer of orbitals. Empty lines and lines starting with {\tt \#, \%, !} are ignored. Notice the swapped indices in the coefficient with respect to the standard 2-body operator formulation $U_{ijkl}c^{\dagger}_{i}c^{\dagger}_{j}c_{l}c_{k}$. To enable reading an interaction matrix file, the flag {\tt ED\_READ\_UMATRIX} needs to be set in the input file. {\tt ED\_READ\_UMATRIX} and {\tt ED\_USE\_KANAMORI} are mutually exclusive, but they can be both false.
Before the impurity problem is solved, the used can also provide interaction terms in the program or script. Extensive documentation is available at (\textcolor{red}{how should we do this?})
All the interaction operators used in each call of the solver are listed and written in an output file. called {\tt UMATRIX\_FILE.used}.

@@

\subsubsection{Sparse map}\label{CodeSparseMap}
As it will be evident in the following \secu{sSecRDM}, there are cases in
which the sector construction requires to store, for any sector state,
disjoint information about the corresponding Fock state.
Let's consider a Fock state $\ket{I}$. Its bit decomposition
$\ket{\vec{n}}$ can be further split into chunks of bit corresponding,
for instance to impurity and bath degrees of freedom as 
$\ket{\vec{n}}=\ket{\vec{i}_\up\,\vec{b}_\up\,\vec{i}_\dw\,\vec{b}_\dw}$.  
The module {\tt ED\_SPARSE\_MAP} implements an hash-table {\tt
  sparse\_map}, which is included as an element in the {\tt sector} object and  
which stores  for every impurity configurations $\vec{i}_\s$ (key) the
bath states $\vec{b}_\s$ (values) corresponding to it.
The {\tt sparse\_map} ${\cal S}$ elements are constructed upon call in {\tt
  build\_sector} for any value of {\tt ed\_mode} using different
algorithms.
In the {\bf normal} case, the sector object contains two {\tt
  sparse\_maps} ${\cal S}_{\s=\up,\dw}$, in line with the
spin dependent factorization of the Fock and sector states: 
$\ket{J}=\ket{J_\up}\otimes\ket{J_\dw} \xleftarrow{\MM}
\ket{j_\up}\otimes\ket{j_\dw}=\ket{j}$.
The sparse maps are build as follows. 
For any spin state $\ket{J_\s}=\ket{{vec{i}_{\s}\vec{b}_{\s}}$ of the sector
the key is determined by the integer $I_\s$ corresponding to the
impurity bit set $\vec{i}_\s$. The values are given by any 
integer $B_\s$ corresponding to any bath bit set
$\vec{b_\s}$ associated with $\vec{i}_\s$.
Thus, any given combination of key-value
reconstructs an integer $J_\s$ representing the Fock states in the
given spin sector according to the rule $J_s = I_\s +
2^{N_{imp}}B_\s$, where $N_{imp}$ is the number of impurity bits.

In the {\bf superc/nonsu2} case the construction of the unique sparse
map is similar. For any Fock state $\ket{J}$ corresponding to sector
state $\ket{j}$ we consider the four integers $I_\up, B_\up, I_\dw,
B_\dw$ corresponding, respectively, to the bit sets
$\vec{i}_\up,\vec{b}_\up,\vec{i}_\dw,\vec{b}_\dw}$ decomposition the
state. These integers fulfill the relation
$
J =I_\up +  B_\up2^{N_{imp}} + (I_\dw +  B_\dw2^{N_{imp}})2^{N_s}
$
To obtain a contiguous memory pattern we define as a key of the sparse
map the combination $I_\up + I_\dw2^{N_{imp}}$ and as corresponding
  values the integer $B_\up + B_\dw2^{N_b}$. Where needed the correct
  bit sets and Fock state can be reconstructed with simple algebra at
  no additional cost.

\setbox0=\hbox{%
  \begin{minipage}{0.5\linewidth}
    \center{\bf normal}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
i=0
do Iup=0,2**Nbit-1     
  nup_ = popcnt(Iup) 
  if(nup_ /= Nups(1))cycle
  i=i+1 
  !...
  iImp  = ibits(iup,0,Norb)
  iBath = ibits(iup,Norb,Norb*Nbath)
  call sp_insert_state(self%H(1)%sp,iImp,iBath,i)
enddo
i=0
do Idw=0,2**Nbit-1
  ndw_= popcnt(Idw)
  if(ndw_ /= Ndws(1))cycle
  i=i+1 
  !...
  iIMP  = ibits(idw,0,Norb)
  iBATH = ibits(idw,Norb,Norb*Nbath)
  call sp_insert_state(self%H(2)%sp,iImp,iBath,dim) 
enddo
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.5\linewidth}
    \center{\bf superc/nonsu2}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
i=0
do Idw=0,2**Ns-1
  ndw_= popcnt(idw)
  do Iup=0,2**Ns-1
    nup_ = popcnt(iup)
    !superc 
    sz_  = nup_ - ndw_
    if(sz_ /= self%Sz)cycle
    !nonsu2 
    !nt_  = nup_ + ndw_
    !if(nt_ /= self%Ntot)cycle 
    i=i+1 
    !...
    iImpUp  = ibits(iup,0,Norb)
    iImpDw  = ibits(idw,0,Norb)
    iBathUp = ibits(iup,Norb,Norb*Nbath)
    iBathDw = ibits(idw,Norb,Norb*Nbath)
    iImp    = iImpUp  + iImpDw*(2**Norb)
    iBath   = iBathUp + iBathDw*(2**(Norb*Nbath))
    call sp_insert_state(self%H(1)%sp,iImp,iBath,dim)
  enddo
enddo

@@

\subsection{Bath parametrization}\label{sSecBath}
The bath parametrization is a key feature in determining the properties
of the quantum impurity problem.
Following the structure of Eq.~\ref{Hbath} the bath can be
parametrized by two terms: the Hamiltonian matrices $h^p$ and the
amplitudes $V^p$, for $p=1,\dots,N_{bath}$.
Internally, the bath is represented by a dedicated object {\tt
  effective\_bath} defined in {\tt ED\_VARS\_GLOBAL}. 
On the
user side all parameters are packed into a rank-1 array of
doubles handled using reverse communication strategy.
This ensures that a local array is always conserved by the user, while
it prevents direct access to the internal copy.

>>>>

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a general quantum impurity problem defined by the
following Hamiltonian:
$$
\hat{H} = \hat{H}_{imp} + \hat{H}_{bath} + \hat{H}_{hyb} + \hat{H}_\mathrm{ph} + \hat{H}_{e-ph}
$$
which describes a  multi-orbital interacting quantum
impurity coupled to an electronic bath and to local, i.e. Holstein,
phonons. 
\begin{equation}\label{Himp}
  \begin{split}
    \hat{H}_{imp} & = \hat{H}^0_{imp} + \hat{H}^{int}_{imp}\\
    \hat{H}^0_{imp} & =
    \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}d_{\b\sigma'}\\
\end{split}
\end{equation}

@@

The bath part and its coupling to the impurity has the form: 
\begin{equation}\label{Hbath}
  \begin{split}
    \hat{H}_{bath} &=
    \sum_p\sum_{\a\b\sigma\sigma'}h^p_{\a\b\sigma\sigma'}a^{+}_{p\a\sigma}a_{p\b\sigma'}\\
    %
    \hat{H}_{hyb} &= \sum_p\sum_{\a\b\sigma\sigma'}V^p_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}a_{p\b\sigma'}+H.c. \\
\end{split}
\end{equation}
where $p=1,\dots,N_\mathrm{bath}$ is an index running over a finite number of
bath elements, $a_{p\alpha\sigma}$ ($a^+_{p\alpha\sigma}$) are the destruction (creation) operators for
the bath electrons with index $p$, with orbital $\alpha$ and
spin $\sigma$.
The  properties of each bath level are described by the
matrices $h^p_{\a\b\sigma\sigma'}$. As such any bath element can be
composed of several electronic levels according the bath topology,
which will be discussed further in the following.
Each bath level couples to the impurity with an amplitude
$V^p_{\a\b\s\sigma'}$ which we allow to couple different orbital and
opposite spins.

Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, $b_q$
($b_q^+$) are the destruction (creation) operators for the phonon $q$
with frequency $\omega_{0q}$. The matrix  $g_{\a\b}$ expresses is the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$. 
In the following we consider a bath discretized into a  number
of bath degrees of freedom and a finite number of available phonons $N_\mathrm{ph}$,
to cut-off the unbounded dimensions of the local phonons Hilbert space.

\paragraph{{\bf Remark}} As extensively discussed in
\onlinecite{Amaricci2022}, the inclusion of local phonons with
truncated dimensions ultimately amounts to a sequential application of
the procedures defined for the electronic part, i.e. one per phonon
mode. This includes a linear scaling of the dimensions with $N_\mathrm{ph}$
and of course largely limit the available degrees of freedom to
describe electronic states. In the rest of the paper we focus
specifically on the electronic part of the quantum impurity problem,
referring to the phonons in specific sections when their presence
introduces non-trivial modifications.

More specifically, we consider a system composed of $N_\mathrm{imp}=1$
impurities, i.e. a single impurity problem, $N_\mathrm{bath}$ bath elements
and $N_\mathrm{ph}$ phonons. The size of the system is determined by
the number of phonons (fixed) and that of {\it electronic} levels, i.e. levels with a local
electronic Hilbert space
$\HH_\mathrm{e}=\{\ket{0},\ket{\up},\ket{\dw},\ket{\up\dw} \}$. 
Due to is internal structure, the single impurity contains
$N_i= N_\a$ electronic levels.
The total number of levels is then determined by
the electronic levels in the bath $N_\mathrm{b}$. This is a function of the bath topology and $N_\mathrm{bath}$,
i.e. the number of bath elements. In the simplest case each bath
element corresponds to an independent electronic levels coupled to the
impurity, thus $N_\mathrm{b}\equiv N_\mathrm{bath}$. 
We indicate with $N_\mathrm{s}=N_i + N_\mathrm{b}$ the total number of electronic levels.

The setup of the general quantum impurity problem is implemented in
different parts of the \NAME software. The dimensions of the system
are controlled by input variables {\tt Nspin}$=N_\sigma$,
{\tt Norb}$=N_\a$ and {\tt Nbath}$=N_\mathrm{bath}$ in {\tt
  ED\_INPUT\_VARS}. These are used to
determine the variables {\tt Ns}$=N_\mathrm{s}$ and $N_\mathrm{b}$, defined in the global
memory pool {\tt ED\_VARS\_GLOBAL}, using the functions contained
in {\tt ED\_SETUP}. 
The user can define the local non-interacting Hamiltonian
$h^0_{\a\b\s\s'}$ using the function {\tt ed\_set\_hloc} in {\tt ED\_AUX\_FUNX}.
The matrix is then stored in the internal memory and shared throughout the
code.
On the other hand the setup of the bath matrices $h^p_{\a\b\s\s'}$
requires a more involved procedure which will be illustrated in
\secu{sSecBath}.

\subsection{The Fock basis states}\label{sSecBasis}
The Fock space of the quantum impurity problems is defined as
$\FF=\FF_\mathrm{e}\otimes \FF_\mathrm{ph}$, with $\FF_\mathrm{e}=\bigoplus_{n=0}^{N_\mathrm{s}}
S_-\HH_\mathrm{e}^{\otimes n}$ the electronic Fock space,  $\FF_\mathrm{ph}=\bigoplus_{n=0}^{N_q}S_+\HH_\mathrm{ph}^{\otimes n}$ the
phonon Fock space,  $\HH_\mathrm{ph}=\{\ket{0},\ket{1},\dots,\ket{N_\mathrm{ph}}\}$ is the local phonon Hilbert space
and ($S_-$)  $S_+$ the (anti-)symmetrization operator.  
The total dimension of the Fock space is
$D=D_\mathrm{e}\cdot D_\mathrm{ph}=4^{N_\mathrm{s}}\cdot (N_\mathrm{ph}+1)$ making the exponential
growth with the number of electron levels transparent. 
The quantum states in the space $\FF$ are naturally represented in
terms of occupation number formalism of the second quantization,
i.e. the Fock basis.
For a system of $N_\mathrm{s}$ electrons each Fock state
reads $\ket{p}\ket{\vec{n}}$ with
\begin{equation}
    \ket{\vec{n}}=\ket{\vec{n}_\up}\ket{\vec{n}_\dw}=\ket{n_{1\up},\dots,n_{N_\mathrm{s}\up},n_{1\dw},\dots,n_{N_\mathrm{s}\dw}},
    \label{eq:FockState_|nupndw>}
\end{equation}
where $p=1,\dots,N_\mathrm{ph}$ is the number of local phonons while $n_{a\sigma}=0,1$ signals the absence or the
presence of an electron with spin $\sigma$ at the level $a$.
The electronic part of the Fock state $\ket{\vec{n}}$ is represented as a string of
zeros and ones of length $2N_\mathrm{s}$. Thus, any such state can  be encoded
in a computer using a sequence of $2N_\mathrm{s}$ bits or, analogously, as a
given integer $I=0,\dots 2^{2N_\mathrm{s}}-1$ so that $\ket{\vec{n}}=\ket{I}$.  
Together with the basis states one defines  destruction and creation 
operators, respectively $c_{a\s}$ and $c^+_{a\s}$, which acts on the
Fock space as: 
$\ket{\vec{n}}$ as:  
\begin{align*}
  c_{a\sigma}\ket{\vec{n}} &=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};\qquad
    c^{+}_{a\sigma}\ket{\vec{n}} &=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
\end{align*}
where $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$ takes care
of the fermionic sign imposed by Pauli principle.

@@

\begin{table}%[ht]
  \label{TabSector}
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_\mathrm{s}}{N_\up}\binom{N_\mathrm{s}}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_\mathrm{s}-S_z-2m}\binom{N_\mathrm{s}}{N_\mathrm{s}-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv N_\up+N_\dw$ & $\binom{2N_\mathrm{s}}{N}$ \\ 
 \hline
\end{tabularx}
\end{center}
\caption{Ciao}
\end{table}

The {\bf normal} case requires a brief remark. Because of the
independent conservation of $N_\up$ and $N_\dw$, the local Hilbert
space and the electronic Fock can be factorizes as, respectively,
$\HH=\HH_\up\otimes\HH_\dw$, $\FF_\mathrm{e} = \FF_{e\up}\otimes \FF_{e\dw}$.  
Accordingly any Fock state is written as $\ket{\vec{n}_\up}\ket{\vec{n}_\dw}$, the symmetry sector can be written as  $\SS_{\vec{Q}} = \SS_{N_\up}\otimes
\SS_{N_\dw}$ and, correspondingly, the sector map splits in two
mutually exclusive ones $\MM = \MM_\up
\otimes \MM_\dw$.
Thus, each state $\ket{i}=\ket{i_\up}\ket{i_\dw}$ of the
sector is labelled by two integers $[i_\up,i_\dw]$, 
$i_\sigma=1,\dots,D_{\SS_\sigma}$ such that $i=i_\up + i_\dw
D_{\SS_\dw}$. The map $\MM$ connects any such state to a Fock state
$\ket{I}=\ket{I_\up}\ket{I_\dw}$ labelled by two integers
$[I_\up,I_\dw]$ as $I=I_\up +   I_\dw 2^{N_\mathrm{s}}$. For a thorough
discussion about Fock basis organization in this case see
\onlinecite{Amaricci2022}.

@@

where {\tt NORB} is the numer of orbitals. Empty lines and lines starting with {\tt \#, \%, !} are ignored. Notice the swapped indices in the coefficient with respect to the standard 2-body operator formulation $U_{ijkl}c^{\dagger}_{i}c^{\dagger}_{j}c_{l}c_{k}$. To enable reading an interaction matrix file, the flag {\tt ED\_READ\_UMATRIX} needs to be set in the input file. {\tt ED\_READ\_UMATRIX} and {\tt ED\_USE\_KANAMORI} are mutually exclusive, but they can be both false.
Before the impurity problem is solved, the user can also provide interaction terms in the program or script. Extensive documentation is available at (\textcolor{red}{how should we do this?})
All the interaction operators used in each call of the solver are listed and written in an output file, called {\tt UMATRIX\_FILE.used}.

@@

%GBtodo placeholder: I passed on this already (except I don't like the minipage of the listings)
\subsubsection{Sparse map}\label{CodeSparseMap}
As it will be evident in the following \secu{sSecRDM}, there are cases in
which the sector construction requires to store, for any sector state,
disjoint information about the corresponding Fock state.
Let's consider a Fock state $\ket{I}$. Its bit decomposition
$\ket{\vec{n}}$ can be further split into arbitrarily grouped chunks of bits.
For instance, we can choose to group the bits into impurity and bath degrees of freedom,
as $\ket{\vec{n}}=\ket{\vec{i}_\up\,\vec{b}_\up\,\vec{i}_\dw\,\vec{b}_\dw}$, where we 
highlight the overall ordering in spin-up and spin-down parts underlying the whole
codebase (see \equ{eq:FockState_|nupndw>} and the subsequent discussion).  
The module {\tt ED\_SPARSE\_MAP} implements an hash-table {\tt sparse\_map}, 
as a component of the {\tt sector} object, which stores for every impurity 
configuration $\vec{i}_\s$ (\emph{key}) all the possible bath states $\vec{b}_\s$ (\emph{values})
that are compatible with the given conserved quantum numbers (see \secu{sSecQNs}).
The {\tt sparse\_map} ${\cal S}$ objects are constructed upon call of the {\tt
  build\_sector} subroutine, automatically dispatched to a different algorithm 
for each value of {\tt ed\_mode}.
In the {\bf normal} case, the sector object contains two {\tt
  sparse\_maps} ${\cal S}_{\s=\up,\dw}$, in line with the
spin dependent factorization of the Fock and sector states: 
$\ket{J}=\ket{J_\up}\otimes\ket{J_\dw} \xleftarrow{\MM}
\ket{j_\up}\otimes\ket{j_\dw}=\ket{j}$.
The sparse maps are built as follows. 
For any spin state of the sector $\ket{J_\s}=\ket{{\vec{i}_{\s}\vec{b}_{\s}}}$,
the \emph{key} is determined by the integer $I_\s$ corresponding to the
impurity bitset $\vec{i}_\s$. The \emph{values} are given by any
integer $B_\s$ corresponding to any bath bitset
$\vec{b_\s}$ associated with $\vec{i}_\s$ by the conservation law of the 
given sector: $\mathtt{sum}(\vec{i}_\s) + \mathtt{sum}(\vec{b}_\s) = \mathtt{popcnt}(I_\s) + \mathtt{popcnt}(B_\s)= N_\s$.
Thus, any given \emph{key-value} combination 
reconstructs an integer $J_\s$ representing a Fock state in the
given spin sector according to the rule $J_s = I_\s +
2^{N_\mathrm{imp}}B_\s$, where $N_\mathrm{imp}$ is the number of impurity bits.

In the {\bf superc/nonsu2} case the construction of the unique sparse
map is similar. For any Fock state $\ket{J}$ corresponding to sector
state $\ket{j}$ we consider the four integers $I_\up, B_\up, I_\dw,
B_\dw$ corresponding, respectively, to the bitset
$\ket{\vec{i}_\up,\vec{b}_\up,\vec{i}_\dw,\vec{b}_\dw}$ decomposition the
state. These integers fulfill the relation
$
J =I_\up +  B_\up2^{N_\mathrm{imp}} + (I_\dw +  B_\dw2^{N_\mathrm{imp}})2^{N_\mathrm{s}}
$
To obtain a contiguous memory pattern we define as a \emph{key} of the sparse
map the combination $I_\up + I_\dw2^{N_\mathrm{imp}}$ and as corresponding
\emph{values} the integers $B_\up + B_\dw2^{N_\mathrm{b}}$, fulfilling the
corresponding conservation laws: 
$\mathtt{popcnt}(I_\up + I_\dw2^{N_\mathrm{imp}}) + \mathtt{popcnt}(B_\up + B_\dw2^{N_\mathrm{b}}) = N_\up + N_\dw$
for the \textbf{nonsu2} case and
$\mathtt{popcnt}(I_\up + B_\up2^{N_\mathrm{imp}}) - \mathtt{popcnt}(I_\dw + B_\dw2^{N_\mathrm{imp}}) = S_z$
for the \textbf{superc} case.
Where needed, the corresponding bitset and Fock state can be reconstructed 
with simple algebra at no additional cost.

\setbox0=\hbox{%
  \begin{minipage}{0.48\linewidth}
    \center{\bf normal}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
i=0
do Iup=0,2**Nbit-1     
  nup_ = popcnt(Iup) 
  if(nup_ /= Nups(1))cycle
  i=i+1 
  !...
  iImp  = ibits(iup,0,Norb)
  iBath = ibits(iup,Norb,Norb*Nbath)
  call sp_insert_state(self%H(1)%sp,iImp,iBath,i)
enddo
i=0
do Idw=0,2**Nbit-1
  ndw_= popcnt(Idw)
  if(ndw_ /= Ndws(1))cycle
  i=i+1 
  !...
  iIMP  = ibits(idw,0,Norb)
  iBATH = ibits(idw,Norb,Norb*Nbath)
  call sp_insert_state(self%H(2)%sp,iImp,iBath,dim) 
enddo
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.52\linewidth}
    \center{\bf superc/nonsu2}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
i=0
do Idw=0,2**Ns-1
  ndw_= popcnt(idw)
  do Iup=0,2**Ns-1
    nup_ = popcnt(iup)
    select case (ed_mode)
      case('superc') 
        sz_  = nup_ - ndw_
        if(sz_ /= self%Sz)cycle
      case('nonsu2') 
        nt_  = nup_ + ndw_
        if(nt_ /= self%Ntot)cycle 
    end select
    i=i+1 
    !...
    iImpUp  = ibits(iup,0,Norb)
    iImpDw  = ibits(idw,0,Norb)
    iBathUp = ibits(iup,Norb,Norb*Nbath)
    iBathDw = ibits(idw,Norb,Norb*Nbath)
    iImp    = iImpUp  + iImpDw*(2**Norb)
    iBath   = iBathUp + iBathDw*(2**(Norb*Nbath))
    call sp_insert_state(self%H(1)%sp,iImp,iBath,dim)
  enddo
enddo

@@

\subsection{Bath parametrization}\label{sSecBath}
The bath parametrization is a key feature in determining the properties
of the quantum impurity problem.
Following the structure of Eq.~\ref{Hbath} the bath can be
parametrized by two terms: the Hamiltonian matrices $h^p$ and the
amplitudes $V^p$, for $p=1,\dots,N_\mathrm{bath}$.
Internally, the bath is represented by a dedicated object {\tt
  effective\_bath} defined in {\tt ED\_VARS\_GLOBAL}. 
On the
user side all parameters are packed into a rank-1 array of
doubles handled using reverse communication strategy.
This ensures that a local array is always conserved by the user, while
it prevents direct access to the internal copy.
